Optimizing Weighted Lower Linear Envelope Potentials Within Latent-SVM Framework

### 2016-08-22 Report/draft.tex
Amended index error
Started Final Report

\DeclareMathOperator*{\argmax}{arg\,max}

@@

	\section{Latent Structural SVMs}
		To include unobserved information in the model, Yu\cite{yu2009learning} extended the joint feature function\cite{tsochantaridis2005large}  $\Psi(\mathbf{x},\mathbf{y}) $ with a latent variable $\mathbf{h}\in \mathcal{H}$ to $\Psi(\mathbf{x},\mathbf{y},\mathbf{h}) $. So the inference problem becomes
		$$
		f_w(x) = \argmax_{(\mathbf{y} \times \mathbf{h}) \in \mathcal{Y} \times \mathcal{H}} w\cdot\Psi(\mathbf{x},\mathbf{y},\mathbf{h})
		$$
		Accordingly, the loss function can be extended as
		$$
		\Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w)))
		$$
		where
		$$
		\mathbf{h}^*_i(w) = \argmax_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})
		$$
		$$
		(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w))=\argmax_{(\mathbf{y} \times \mathbf{h}) \in \mathcal{Y} \times \mathcal{H}} w\cdot\Psi(\mathbf{x}_i,\mathbf{y},\mathbf{h})
		$$
		However, under this formulation the ``loss augmented inference'' used in Structral SVMs\cite{tsochantaridis2005large} to remove the complexity cannot be performed due to the dependence of loss function $\Delta$ on hidden variables $\mathbf{h}^*_i(w)$. Yu argued that in real world applications hidden variables are usually intermediate results and are not required as an output\cite{yu2009learning}. Therefore, the loss function can only focus on the inferenced hidden variables $\mathbf{\hat{h}}_i(w)$ which leads to:
		$$
		\Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w))) = \Delta(\mathbf{y}_i,\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w))
		$$
		Thus the upper bound used in structural SVMs can be written as:
		\begin{align*}
			\Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w)))
			&\leq \bigg(\max_{(\mathbf{\hat{y}} \times \mathbf{\hat{h}}) \in \mathcal{Y} \times \mathcal{H}} [w\cdot\Psi(\mathbf{x}_i,\mathbf{\hat{y}},\mathbf{\hat{h}}) + \Delta(\mathbf{y}_i,\mathbf{\hat{y}},\mathbf{\hat{h}})]\bigg)\\
			&-\max_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})
		\end{align*}
		Hence the optimization problem for Structural SVMs with latent variables becomes
		\begin{align*}
			\min_w\bigg(\frac{1}{2}\|w\|^2+
			C\sum_{i=1}^{n}\big(\max_{(\mathbf{\hat{y}} \times \mathbf{\hat{h}}) \in \mathcal{Y} \times \mathcal{H}} [w\cdot\Psi(\mathbf{x}_i,\mathbf{\hat{y}},\mathbf{\hat{h}}) + \Delta(\mathbf{y}_i,\mathbf{\hat{y}},\mathbf{\hat{h}})]\big)\bigg)\\
			-C\sum_{i=1}^{n}\big(\max_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})\big)
		\end{align*}
		Observing that the former equation is a difference of two convex functions, therefore the redefined objective function can be optimized locally using CCCP algorithm.

@@

		\section{Optimizing Lower Linear Envelope Using Latent Structural SVMs}
		To optimize the energy function using latent structral SVMs, it first requires us write the equation (5) into a linear combination of weights and feature vector. In order to help with the formulation, we first define $k^* = \argmax_{k=1,\dots,K-1} \big[\big[z_{k}\big]\big]$. We also notice that the equation (5) can be rewrite as an inner product of two vectors ($b_1=0$ according to equation  (4)):
		\begin{align*}
		E^c(y_c,z)&=a_1W_c(y_c)+b_1+\sum_{k=1}^{K-1}z_k((a_{k+1}-a_k)W_c(y_c)+b_{k+1}-b_k)\\
		&=a_1W_c(y_c)+\sum_{k=1}^{K-1}(a_{k+1}-a_k)z_kW_c(y_c)+\sum_{k=1}^{K-1}(b_{k+1}-b_k)z_k
		\end{align*}
		One thing to notice is that $z_{k^*+1}=0$. From equation (6) we know that $z_k = 1$ from $k=1$ to $k=k^*$ and all variables $z_k$ after $k^*$ are $0$s. We can reparameterized the energy function as $E^c(y_c,z)=\mathbf{w}\mathbf{\phi}$ where:
		\begin{equation}
			w_i = \left\{
			\begin{aligned}
				& a_1	& \text{for} \ i=1\\
				& a_i-a_{i-1} & \text{for}\ 1< i \leq K\\
				& b_{i+1-K}-b_{i-K} & \text{for} \ K<i\le2K-1\\
			\end{aligned}
			\right.
		\end{equation}

		\begin{equation}
		\phi_i = \left\{
		\begin{aligned}
		& W(\mathbf{y}) 	& \text{for} \ i=1\\
		& W(\mathbf{y})\bigg[\bigg[i\le k^*\bigg]\bigg] & \text{for}\ 1<i\le i\\
		& \bigg[\bigg[ i\le k^*\bigg]\bigg]  & \text{for} \ K<i\le2K-1\\
		\end{aligned}
		\right.
		\end{equation}
		Thus our inference problem becomes:
		$$
		(\mathbf{\hat{y}}_i(w),\mathbf{\hat{z}}_i(w))=\argmax_{(\mathbf{y} \times \mathbf{z}) \in \mathcal{Y} \times \mathcal{Z}} w\cdot\Psi(\mathbf{y},\mathbf{z})
		$$
		$$
		\mathbf{z}^*_i(w) = \argmax_{\mathbf{z} \in \mathcal{Z}} w \cdot \Psi(\mathbf{y}_i,\mathbf{z})
		$$
		We can perform the graph-cuts algorithm proposed by Gould\cite{gouldlearning} on both equations. For the later one we simply use the ground-truth labels $y_i$ as function input. Together with linear constraints (2) and (3), we can optimize the energy function using latent SVMs.

>>>>

% Algorithm Packages
% ---
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\makeatletter

% Defines
% ---
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

	\section{Latent Structural SVMs}
		To include unobserved information in the model, Yu\cite{yu2009learning} extended the joint feature function\cite{tsochantaridis2005large}  $\Psi(\mathbf{x},\mathbf{y}) $ with a latent variable $\mathbf{h}\in \mathcal{H}$ to $\Psi(\mathbf{x},\mathbf{y},\mathbf{h}) $. So the inference problem becomes
		$$
		f_w(x) = \argmax_{(\mathbf{y} \times \mathbf{h}) \in \mathcal{Y} \times \mathcal{H}} w\cdot\Psi(\mathbf{x},\mathbf{y},\mathbf{h})
		$$
		Accordingly, the loss function can be extended as
		$$
		\Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w)))
		$$
		where
		$$
		\mathbf{h}^*_i(w) = \argmax_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})
		$$
		$$
		(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w))=\argmax_{(\mathbf{y} \times \mathbf{h}) \in \mathcal{Y} \times \mathcal{H}} w\cdot\Psi(\mathbf{x}_i,\mathbf{y},\mathbf{h})
		$$
		However, under this formulation the ``loss augmented inference'' used in Structral SVMs\cite{tsochantaridis2005large} to remove the complexity cannot be performed due to the dependence of loss function $\Delta$ on hidden variables $\mathbf{h}^*_i(w)$. Yu argued that in real world applications hidden variables are usually intermediate results and are not required as an output\cite{yu2009learning}. Therefore, the loss function can only focus on the inferenced hidden variables $\mathbf{\hat{h}}_i(w)$ which leads to:
		$$
		\Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w))) = \Delta(\mathbf{y}_i,\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w))
		$$
		Thus the upper bound used in structural SVMs can be written as:
		\begin{align}
			\label{eq:upperbound_SSVM}
			\Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w)))
			&\leq \bigg(\max_{(\mathbf{\hat{y}} \times \mathbf{\hat{h}}) \in \mathcal{Y} \times \mathcal{H}} [w\cdot\Psi(\mathbf{x}_i,\mathbf{\hat{y}},\mathbf{\hat{h}}) + \Delta(\mathbf{y}_i,\mathbf{\hat{y}},\mathbf{\hat{h}})]\bigg)\\
			&-\max_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})
		\end{align}
		Hence the optimization problem for Structural SVMs with latent variables becomes
		\begin{align*}
			\min_w\bigg(\frac{1}{2}\|w\|^2+
			C\sum_{i=1}^{n}\big(\max_{(\mathbf{\hat{y}} \times \mathbf{\hat{h}}) \in \mathcal{Y} \times \mathcal{H}} [w\cdot\Psi(\mathbf{x}_i,\mathbf{\hat{y}},\mathbf{\hat{h}}) + \Delta(\mathbf{y}_i,\mathbf{\hat{y}},\mathbf{\hat{h}})]\big)\bigg)\\
			-C\sum_{i=1}^{n}\big(\max_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})\big)
		\end{align*}
		Observing that the former equation is a difference of two convex functions, therefore the redefined objective function can be optimized locally using CCCP algorithm.

@@

		\section{Optimizing Lower Linear Envelope Using Latent Structural SVMs}
		To optimize the energy function using latent structral SVMs, it first requires us write the equation (5) into a linear combination of weights and feature vector. In order to help with the formulation, we first define $k^* = \argmax_{k=1,\dots,K-1} \big[\big[z_{k}\big]\big]$. We also notice that the equation (5) can be rewrite as an inner product of two vectors ($b_1=0$ according to equation  (4)):
		\begin{align}
		\label{eq:originalenergy}
		E^c(y_c,z)&=a_1W_c(y_c)+b_1+\sum_{k=1}^{K-1}z_k((a_{k+1}-a_k)W_c(y_c)+b_{k+1}-b_k)\\
		&=a_1W_c(y_c)+\sum_{k=1}^{K-1}(a_{k+1}-a_k)z_kW_c(y_c)+\sum_{k=1}^{K-1}(b_{k+1}-b_k)z_k
		\end{align}
		One thing to notice is that $z_{k^*+1}=0$. From equation (6) we know that $z_k = 1$ from $k=1$ to $k=k^*$ and all variables $z_k$ after $k^*$ are $0$s. We can reparameterized the energy function as $E^c(y_c,z)=\mathbf{\theta}\mathbf{\phi}$ where:
		\begin{equation}
			\theta_i = \left\{
			\begin{aligned}
				& a_1	& \text{for} \ i=1\\
				& a_i-a_{i-1} & \text{for}\ 1< i \leq K\\
				& b_{i+1-K}-b_{i-K} & \text{for} \ K<i\le2K-1\\
			\end{aligned}
			\right.
		\end{equation}

		\begin{equation}
		\phi_i = \left\{
		\begin{aligned}
		& W(\mathbf{y}) 	& \text{for} \ i=1\\
		& W(\mathbf{y})\bigg[\bigg[i-1\le k^*\bigg]\bigg] & \text{for}\ 1<i\le K\\
		& \bigg[\bigg[ i-K\le k^*\bigg]\bigg]  & \text{for} \ K<i\le2K-1\\
		\end{aligned}
		\right.
		\end{equation}
		However, notice that by using energy function labels are infered by minimization. Thus our inference problem becomes:
		$$
		(\mathbf{\hat{y}}_i(\theta),\mathbf{\hat{z}}_i(\theta))=\argmin_{(\mathbf{y} \times \mathbf{z}) \in \mathcal{Y} \times \mathcal{Z}} \theta\cdot\Psi(\mathbf{y},\mathbf{z})
		$$
		$$
		\mathbf{z}^*_i(\theta) = \argmin_{\mathbf{z} \in \mathcal{Z}} \theta \cdot \Psi(\mathbf{y}_i,\mathbf{z})
		$$
		We can perform the graph-cuts algorithm proposed by Gould\cite{gouldlearning} on both equations. For the later one we simply use the ground-truth labels $y_i$ as function input. Thus the term $((a_{k+1}-a_k)W_c(y_c)+b_{k+1}-b_k)$ in equation~\ref{eq:originalenergy} becomes constant, so we can infer latent variable $z$ explicitly by:
		\begin{align}
		\label{eq:inferlatent}
		z_k^c &=
		\begin{cases}
		0 & \text{if $((a_{k+1}-a_k)W_c(y_c)+b_{k+1}-b_k)>=0$} \\
		1 & \text{otherwise}.
		\end{cases}
		\end{align}

		Therefore, the upperbound of equation~\ref{eq:upperbound_SSVM} under this formulation becomes:
		\begin{align}
		\label{eq:upperbound_LSSVM}
		\Delta((\mathbf{y}_i,\mathbf{z}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{z}}_i(w)))
		&\leq \bigg(\min_{(\mathbf{\hat{y}} \times \mathbf{\hat{z}}) \in \mathcal{Y} \times \mathcal{Z}} [w\cdot\Psi(\mathbf{x}_i,\mathbf{\hat{y}},\mathbf{\hat{z}}) - \Delta(\mathbf{y}_i,\mathbf{\hat{y}},\mathbf{\hat{z}})]\bigg)\\
		&-\min_{\mathbf{z} \in \mathcal{Z}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{z})
		\end{align}
		Therefore, the energy function (higher order features together with unary and pairwise features) are:
			\begin{equation}
			E^{all}(y,z) = \begin{bmatrix}
			\theta^h\\
			\theta^{unary}\\
			\theta^{pairwise}
			\end{bmatrix}^T 
			\cdot \begin{bmatrix}
			\phi^h\\
			\phi^{unary}\\
			\phi^{pairwise}
			\end{bmatrix}=\theta^{all\text{T}}\cdot\phi^{all}
			\end{equation}
		Together with linear constraints (2) and (3), we can optimize the energy function using latent SVMs:
		\begin{align}
		\label{eq:qpobject}
		\text{minimize}    \frac{1}{2}||\theta||^2 + \frac{C}{T}\sum_{t=1}^T{\xi_t} \\
		\text{subject to}\\
				\bigg(\min_{(\mathbf{\hat{y}} \times \mathbf{\hat{h}}) \in \mathcal{Y} \times \mathcal{H}} [w\cdot\Psi(\mathbf{x}_i,\mathbf{\hat{y}},\mathbf{\hat{h}}) - \Delta(\mathbf{y}_i,\mathbf{\hat{y}},\mathbf{\hat{h}})]\bigg)
				&-\min_{\mathbf{h} \in \mathcal{H}} w \cdot \Psi(\mathbf{x}_i,\mathbf{y}_i,\mathbf{h})
				+\xi_i\\
				&\geq \Delta((\mathbf{y}_i,\mathbf{h}^*_i(w)),(\mathbf{\hat{y}}_i(w),\mathbf{\hat{h}}_i(w)))
		\end{align}

	\section{Algorithm}

		\begin{algorithm}
							\caption{MRF-LSSVM (CCCP-Cutting Plane)}\label{euclid}
							\begin{algorithmic}[1]
								\State Initialize $\theta=0$
								\State \emph{Outer Loop}:
								\While{t $\leq$ MaxOuterIteration}
								\State Infer latent variable $z^*$ using equation~\ref{eq:inferlatent}
								\State $v_t = \phi(y, z^*)$
								\State \emph{Innter Loop}:
								\State $A \gets$ convex constraints
								\State $b \gets$ convex constraints
								\State $\Delta_{unaryloss} \gets y - \bar{y}$
								\While{t $\leq$ MaxInnerIteration}
								\State $y_{loss}, z_{loss} \gets \argmin_{(\mathbf{y} \times \mathbf{z}) \in \mathcal{Y} \times \mathcal{Z}}  \theta \cdot \Psi(\mathbf{y} - \Delta_{unaryloss},z)$
								\State $\phi_{loss} \gets \phi(x,y_{loss}, z_{loss})$
								\State $A \cup v_t - \phi_{loss}$
								\State $b \cup \Delta(y,y_{loss})$
								\EndWhile
								\EndWhile
							\end{algorithmic}
		\end{algorithm}
